
```{r }
pacman::p_load(data.table, dplyr, ggplot2, tidyr)

base_folder = "C:/models/treso-ldpm/output"
```

This report summarizes the comparison between runs of the LDPM model. The first group corresponds to 20 different
random seeds and is called "with_different_seeds". The second one corresponds to just three replications of the random
seed 1. Additionally, these last three replications were ran withpout paralellizing streams in java (i.e. use forEachOrdered() 
methods instead of forEach()). This keeps the order of sequences of elements in a loop as deterministic; the runtime is 
multiplied by four (THIS IS NOT IMPEMENTED IN THE CURRENT TRESO-LDPM VERSION).  

### Reading the data

```{r }
seeds = c("with_same_seed", "with_different_seeds")

#We run 20 iterations
iterations = 0:19

#Read all the data and summarize it
trips_by_zone = data.frame()
trips_by_ldpm_zone = data.frame()
trips_by_dest_ldpm_zone = data.frame()
trips_by_ldpm_zone_and_mode = data.frame()
trips_by_dest_zone = data.frame()


for (seed in seeds){
  for (iteration in iterations){
    file_name = paste(base_folder, seed, iteration,"ldpm_trips.csv", sep = "/" )
    data = try(fread(file_name),F)
    
    if (is.data.frame(data)){
      #process data and store summaries
      this_trips_by_zone = data %>%
        filter ( tripOriginType == "ONTARIO") %>%
        group_by(tripOriginZone) %>%
        summarize(count = n())
      
      this_trips_by_zone$case = seed
      this_trips_by_zone$iteration = iteration
      this_trips_by_zone$unique_run = paste(seed, iteration, sep = "-")
      trips_by_zone = rbind(trips_by_zone, this_trips_by_zone)
      
      this_trips_by_ldpm_zone = data %>%
        filter ( tripOriginType == "ONTARIO") %>%
        group_by(tripOriginCombinedZone) %>%
        summarize(count = n())
      
      this_trips_by_ldpm_zone$case = seed
      this_trips_by_ldpm_zone$iteration = iteration
      this_trips_by_ldpm_zone$unique_run = paste(seed, iteration, sep = "-")
      trips_by_ldpm_zone = rbind(trips_by_ldpm_zone,this_trips_by_ldpm_zone)
      
      this_trips_by_dest_ldpm_zone = data %>%
        filter ( destZoneType == "ONTARIO") %>%
        group_by(tripDestCombinedZone) %>%
        summarize(count = n())
      
      this_trips_by_dest_ldpm_zone$case = seed
      this_trips_by_dest_ldpm_zone$iteration = iteration
      this_trips_by_dest_ldpm_zone$unique_run = paste(seed, iteration, sep = "-")
      trips_by_dest_ldpm_zone = rbind(trips_by_dest_ldpm_zone,this_trips_by_dest_ldpm_zone)
      
      this_trips_by_ldpm_zone_and_mode = data %>%
        filter ( tripOriginType == "ONTARIO") %>%
        group_by(tripOriginCombinedZone, tripMode) %>%
        summarize(count = n())
      
      this_trips_by_ldpm_zone_and_mode$case = seed
      this_trips_by_ldpm_zone_and_mode$iteration = iteration
      this_trips_by_ldpm_zone_and_mode$unique_run = paste(seed, iteration, sep = "-")
      trips_by_ldpm_zone_and_mode = rbind(trips_by_ldpm_zone_and_mode,as.data.frame(this_trips_by_ldpm_zone_and_mode))
      
      this_trips_by_dest_zone = data %>%
        filter ( destZoneType == "ONTARIO") %>%
        group_by(destZone) %>%
        summarize(count = n())
      
      this_trips_by_dest_zone$case = seed
      this_trips_by_dest_zone$iteration = iteration
      this_trips_by_dest_zone$unique_run = paste(seed, iteration, sep = "-")
      trips_by_dest_zone = rbind(trips_by_dest_zone,as.data.frame(this_trips_by_dest_zone))
    }
  }
}
```

### Number of trips by origin treso zone

We take a subsample of 15 zones for visualization purposes.

```{r }
tripOriginZones = unique(trips_by_zone$tripOriginZone)
tripOriginZonesSample = sample(x = tripOriginZones, size = 15, replace = F)

trips_by_zone_sample = trips_by_zone %>% filter(tripOriginZone %in% tripOriginZonesSample)

averages = trips_by_zone_sample %>% group_by(tripOriginZone, case) %>% summarize(mean = mean(count))

trips_by_zone_sample = merge(trips_by_zone_sample, averages, by = c("tripOriginZone", "case"))


p1 = ggplot(trips_by_zone_sample, aes(x = as.factor(tripOriginZone), group=tripOriginZone, y = (count - mean)/mean*100)) +
  stat_boxplot() + facet_grid(case ~ .) + xlab("zones (subset of 15)") + ylab("differences (%)")

print(p1)
```

Plot p1 above shows the distribution of the difference in the number of trips for each zone with respect of the 
average = number of trips in this replication - average number of trips in all replications. The numbers are expressed in
percentage. The plot shows the box-plot, where the boxes contain the 50% central quantile of the points. 

```{r }
p2 = ggplot(trips_by_zone_sample, aes(x = mean, y = (count - mean)/mean*100)) +
  geom_point(alpha = .1) + facet_grid(case ~ .) + xlab("mean") + ylab("differences (%)")

print(p2)
```

Plot p1 above shows the same variable, ploted agains the mean value of number of trips from the zone. The plot shows that 
the random differences are higher when the average number of trips from the zone is small (at smaller zones). 

```{r }
#Number of trips by origin at ldpm zone

tripOriginCombinedZones = unique(trips_by_ldpm_zone$tripOriginCombinedZone)
tripOriginCombinedZonesSample = sample(x = tripOriginCombinedZones, size = 15, replace = F)

trips_by_ldpm_zone_sample = trips_by_ldpm_zone %>% filter(tripOriginCombinedZone %in% tripOriginCombinedZonesSample)

##get the average values
averages = trips_by_ldpm_zone_sample %>% group_by(tripOriginCombinedZone, case) %>% summarize(mean = mean(count))

trips_by_ldpm_zone_sample = merge(trips_by_ldpm_zone_sample, averages, by = c("tripOriginCombinedZone", "case"))

p3 = ggplot(trips_by_ldpm_zone_sample, aes(x = as.factor(tripOriginCombinedZone), group=tripOriginCombinedZone, y = (count - mean)/mean*100)) +
  stat_boxplot() + facet_grid(case ~ .) + xlab("zones (subset of 15)") + ylab("differences (%)")

print(p3)

p4 = ggplot(trips_by_ldpm_zone_sample, aes(x = mean, y = (count - mean)/mean*100)) +
  geom_point(alpha = .1) + facet_grid(case ~ .) + xlab("mean") + ylab("differences (%)")

print(p4)
```

Plots p3 and p4 refer to the same as p1 and p2, but measured at the LDPM zone resolution. As these zones are larger, the 
random effects are smaller. 

```{r }
#Number of trips by destination at ldpm zone

tripDestCombinedZones = unique(trips_by_dest_ldpm_zone$tripDestCombinedZone)
tripDestCombinedZonesSample = sample(x = tripDestCombinedZones, size = 15, replace = F)

trips_by_dest_ldpm_zone_sample = trips_by_dest_ldpm_zone %>% filter(tripDestCombinedZone %in% tripDestCombinedZonesSample)

averages = trips_by_dest_ldpm_zone_sample %>% group_by(tripDestCombinedZone, case) %>% summarize(mean = mean(count))

trips_by_dest_ldpm_zone_sample = merge(trips_by_dest_ldpm_zone_sample, averages, by = c("tripDestCombinedZone", "case"))

p5 = ggplot(trips_by_dest_ldpm_zone_sample, aes(x = as.factor(tripDestCombinedZone), group=tripDestCombinedZone, y = (count - mean)/mean*100)) +
  stat_boxplot() + facet_grid(case ~ .) + xlab("zones (subset of 15)") + ylab("differences (%)")

print(p5)

p6 = ggplot(trips_by_dest_ldpm_zone_sample, aes(x = mean, y = (count - mean)/mean*100)) +
  geom_point(alpha = .1) + facet_grid(case ~ .) + xlab("mean") + ylab("differences (%)")

print(p6)
```

Plots p5 and p6 refer to the same as p1 and p2, p3, and p4, but measuring the number of tirps by destination
LDPM zone resolution. 

```{r }
#number/ share of trips by mode by origin LDPM zone

tripOriginCombinedZones = unique(trips_by_ldpm_zone_and_mode$tripOriginCombinedZone)
tripOriginCombinedZonesSample = sample(x = tripOriginCombinedZones, size = 10, replace = F)

trips_by_ldpm_zone_and_mode_sample = trips_by_ldpm_zone_and_mode %>%
  filter(tripOriginCombinedZone %in% tripOriginCombinedZonesSample)

###for visualization only - fill the interations missing for the same seed case, if neeed. 

highest_iter = max(trips_by_ldpm_zone_and_mode_sample$iteration)
highest_iter_same_seed = max(trips_by_ldpm_zone_and_mode_sample %>% filter(case == "with_same_seed") %>% select(iteration))
data_for_first_iter = trips_by_ldpm_zone_and_mode_sample %>% filter(case == "with_same_seed", iteration == 1)

while (highest_iter_same_seed < highest_iter){
  highest_iter_same_seed = highest_iter_same_seed + 1; 
  data_for_first_iter$iteration = highest_iter_same_seed
  trips_by_ldpm_zone_and_mode_sample = rbind(trips_by_ldpm_zone_and_mode_sample, data_for_first_iter)
}
  


p7 = ggplot(trips_by_ldpm_zone_and_mode_sample, aes(x = iteration, fill = as.factor(tripMode), weight = count, width=0)) +
  geom_bar(position = "stack", size = 0) + facet_grid(case ~ tripOriginCombinedZone, scale = "free_x") + 
  xlab("zones (subset of 15)") + ylab("trips by mode")

print(p7)
```

Plot p7 shows the absolute number of trips by mode by random seed. Colors refer to mode, where 0 is auto, 1
 is air, 2 is rail and 3 is bus.

```{r }
p8 = ggplot(trips_by_ldpm_zone_and_mode_sample, aes(x = iteration, fill = as.factor(tripMode), weight = count, width=0)) +
  geom_bar(position = "fill", size = 0) + facet_grid(case ~ tripOriginCombinedZone, scale = "free_x") +
  xlab("zones (subset of 15)") +
  ylab("modal share (%)")

print(p8)
```

Plot p8 shows the same variable as p7 but expressend in terms of modal share.

```{r }
#number of trips by destination at TRESO zone

destZones = unique(trips_by_dest_zone$destZone)
destZonesSample = sample(x = destZones, size = 15, replace = F)

trips_by_dest_zone_sample = trips_by_dest_zone %>% filter(destZone %in% destZonesSample)

averages = trips_by_dest_zone_sample %>% group_by(destZone, case) %>% summarize(mean = mean(count))

trips_by_dest_zone_sample = merge(trips_by_dest_zone_sample, averages, by = c("destZone", "case"))

p9 = ggplot(trips_by_dest_zone_sample, aes(x = as.factor(destZone), group=destZone, y = (count - mean)/mean*100)) +
  stat_boxplot() + facet_grid(case ~ .) + xlab("zones (subset of 15)") + ylab("differences (%)")

print(p9)

p10 = ggplot(trips_by_dest_zone_sample, aes(x = mean, y = (count - mean)/mean*100)) +
  geom_point(alpha = .1) + facet_grid(case ~ .) + xlab("mean") + ylab("differences (%)")

print(p10)
```

Plot p9 and p10 show the number of trips by TRESO zone at destination.

---
title: "analyze_random_effects.R"
author: "carlloga"
date: "Fri May 24 14:59:56 2019"
---
